---
title: "2025 Healthcare Survey: Mode Comparison Subgroup Analysis"
author: "Feiran GE & Ruisi Ma"
date: "2026-03-20"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Setup

```{r libraries}
# [MERGED] Combined all libraries from both files into one chunk
# readxl: Feiran | readr removed (replaced by readxl) | broom: moved up from mid-file
library(readxl)
library(dplyr)
library(stringr)
library(tidyr)
library(broom)
```

```{r read-data}
# [MERGED] Single data read for both analyses
# Using read_excel (Feiran's method) since files are .xlsx
# Ruisi's read_csv calls and absolute file paths have been removed
full <- read_excel("/Users/ruisima/Desktop/2025 SLPs Health Care Full Sample for Capstone.xlsx")
resp <- read_excel("/Users/ruisima/Desktop/2025 SLPs Health Care Respondent File for Capstone.xlsx")

# Check column names
names(full)
names(resp)
```

```{r na-check}
# check NA (Feiran)
na_summary_full <- full %>%
  summarise(across(everything(), ~ mean(is.na(.)))) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "na_prop"
  ) %>%
  arrange(desc(na_prop))

na_summary_resp <- resp %>%
  summarise(across(everything(), ~ mean(is.na(.)))) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "na_prop"
  ) %>%
  arrange(desc(na_prop))

na_summary_full; na_summary_resp
```

```{r response-indicator}
# Create response indicator
# 1 = responded, 0 = not responded
full <- full %>% mutate(Response = ifelse(NewID %in% resp$NewID, 1, 0))
```

```{r recode-mode}
# Recode survey mode
full <- full %>%
  mutate(
    Mode = case_when(
      Mode == 1 ~ "Mail",
      Mode == 2 ~ "Web",
      TRUE ~ NA_character_
    )
  )
```

```{r extract-codes}
# [MERGED] Combined code extraction from both files into one mutate()
# facil_code & state_code: Feiran | emp_code & func_code: Ruisi
full <- full %>%
  mutate(
    facil_code = str_extract(Facility, "^\\d{2}"),
    state_code  = str_extract(State, "^\\d{4}"),
    emp_code    = str_extract(Employment_status, "^\\d{2}"),
    func_code   = str_extract(Function, "^\\d{4}")
  )
```

---

# Part 1: Facility Type (Feiran)

```{r recode-facility}
# Recode Facility
# Missing values are kept as NA and excluded only
# when analyzing this subgroup variable

full <- full %>%
  mutate(
    facil_recode = case_when(

      # 1 = hospital-related
      str_detect(Facility, "301|306|307|308") ~ 1,

      # 2 = home health
      str_detect(Facility, "501") ~ 2,

      # 3 = outpatient clinic / office
      str_detect(Facility, "503|504|505|506|510|513") ~ 3,

      # 4 = pediatric hospital
      str_detect(Facility, "304") ~ 4,

      # 5 = rehab hospital
      str_detect(Facility, "303") ~ 5,

      # 6 = SNF
      str_detect(Facility, "401|410") ~ 6,

      # 7 = other
      TRUE ~ 7
    )
  )
```

```{r rr-facility}
# Response rate by Facility subgroup
# Exclude missing Facility only here

rr_facility <- full %>%
  filter(!is.na(facil_recode), !is.na(Mode)) %>%
  group_by(facil_recode, Mode) %>%
  summarise(
    n = n(),
    r = sum(Response, na.rm = TRUE),
    RR = r / n,
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = Mode,
    values_from = c(n, r, RR)
  ) %>%
  mutate(
    Delta = RR_Mail - RR_Web,
    Facility = case_when(
      facil_recode == 1 ~ "General medical / VA / university / military / LTAC hospital",
      facil_recode == 2 ~ "Home health agency / client's home",
      facil_recode == 3 ~ "Outpatient clinic or office",
      facil_recode == 4 ~ "Pediatric hospital",
      facil_recode == 5 ~ "Rehabilitation hospital",
      facil_recode == 6 ~ "Skilled nursing facility",
      facil_recode == 7 ~ "Other",
      TRUE ~ NA_character_
    )
  ) %>%
  relocate(Facility, .before = facil_recode)

rr_facility
```

```{r chisq-facility}
# chi-square for Facility subgroup
chisq_facility <- full %>%
  filter(!is.na(facil_recode), !is.na(Mode), !is.na(Response)) %>%
  group_by(facil_recode) %>%
  group_modify(~{
    tab <- table(.x$Mode, .x$Response)
    test <- chisq.test(tab, correct = FALSE)
    tibble(
      chi_square = unname(test$statistic),
      df = unname(test$parameter),
      p_value = test$p.value
    )
  })
chisq_facility
```

Chi-square tests indicate a statistically significant association between survey mode and response status within all facility types (all p-values < 0.001). This suggests that the effect of mode on response rates is consistent across different work settings, including hospitals, outpatient clinics, home health, and skilled nursing facilities. Importantly, these results imply that the observed mode differences are not driven solely by variation in facility composition, but reflect a persistent within-group mode effect.

```{r models-facility}
# Facility logistic regression
full <- full %>%
  mutate(
    Mode = factor(Mode, levels = c("Web", "Mail")),
    facil_recode = factor(facil_recode)
  )

model_fac_main <- glm(
  Response ~ Mode + facil_recode,
  data = full %>% filter(!is.na(facil_recode), !is.na(Mode)),
  family = binomial()
)

model_fac_int <- glm(
  Response ~ Mode * facil_recode,
  data = full %>% filter(!is.na(facil_recode), !is.na(Mode)),
  family = binomial()
)

lrt_fac <- anova(model_fac_main, model_fac_int, test = "Chisq")
lrt_fac
```

The interaction between mode and facility type is statistically significant (p = 0.024), suggesting that the magnitude of the mode effect varies across facility settings.

```{r summary-facility}
summary(model_fac_int)
```

```{r or-facility}
# odds ratio
tidy(model_fac_int, exponentiate = TRUE, conf.int = FALSE)
```

---

# Part 2: Geographic Region / State (Feiran)

```{r recode-state}
# Recode State for both files
# full: turn State abbreviation to numbers
state_lookup <- c(state.abb, "DC")

state_lookup <- append(state.abb, "DC", after = 8)

full <- full %>%
  mutate(
    state_clean = toupper(State),
    state_code = match(state_clean, state_lookup)
  )

# 9 and 4 Census Regions
full <- full %>%
  mutate(
    states_9 = case_when(
      state_clean %in% c("CT","ME","MA","NH","RI","VT") ~ "New England",
      state_clean %in% c("NJ","NY","PA") ~ "Middle Atlantic",
      state_clean %in% c("IL","IN","MI","OH","WI") ~ "East North Central",
      state_clean %in% c("IA","KS","MN","MO","NE","ND","SD") ~ "West North Central",
      state_clean %in% c("DE","DC","FL","GA","MD","NC","SC","VA","WV") ~ "South Atlantic",
      state_clean %in% c("AL","KY","MS","TN") ~ "East South Central",
      state_clean %in% c("AR","LA","OK","TX") ~ "West South Central",
      state_clean %in% c("AZ","CO","ID","MT","NV","NM","UT","WY") ~ "Mountain",
      state_clean %in% c("AK","CA","HI","OR","WA") ~ "Pacific",
      TRUE ~ NA_character_
    ),

    states_4 = case_when(
      states_9 %in% c("New England","Middle Atlantic") ~ "Northeast",
      states_9 %in% c("East North Central","West North Central") ~ "Midwest",
      states_9 %in% c("South Atlantic","East South Central","West South Central") ~ "South",
      states_9 %in% c("Mountain","Pacific") ~ "West",
      TRUE ~ NA_character_
    )
  )
```

```{r rr-state}
# Response rate by State subgroup

rr_state4 <- full %>%
  filter(!is.na(states_4), !is.na(Mode)) %>%
  group_by(states_4, Mode) %>%
  summarise(
    n = n(),
    r = sum(Response, na.rm = TRUE),
    RR = r / n,
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = Mode,
    values_from = c(n, r, RR)
  ) %>%
  mutate(
    Delta = RR_Mail - RR_Web
  )

rr_state4

rr_state9 <- full %>%
  filter(!is.na(states_9), !is.na(Mode)) %>%
  group_by(states_9, Mode) %>%
  summarise(
    n = n(),
    r = sum(Response, na.rm = TRUE),
    RR = r / n,
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = Mode,
    values_from = c(n, r, RR)
  ) %>%
  mutate(
    Delta = RR_Mail - RR_Web
  )
rr_state9
```

```{r chisq-state}
# chi-square for States
chisq_state4 <- full %>%
  filter(!is.na(states_4), !is.na(Mode), !is.na(Response)) %>%
  group_by(states_4) %>%
  group_modify(~{
    tab <- table(.x$Mode, .x$Response)
    test <- chisq.test(tab, correct = FALSE)
    tibble(
      chi_square = unname(test$statistic),
      df = unname(test$parameter),
      p_value = test$p.value
    )
  })
chisq_state4

chisq_state9 <- full %>%
  filter(!is.na(states_9), !is.na(Mode), !is.na(Response)) %>%
  group_by(states_9) %>%
  group_modify(~{
    tab <- table(.x$Mode, .x$Response)
    test <- chisq.test(tab, correct = FALSE)
    tibble(
      chi_square = unname(test$statistic),
      df = unname(test$parameter),
      p_value = test$p.value
    )
  })

chisq_state9
```

Across all four Census regions, chi-square tests show a significant relationship between survey mode and response status (all p-values < 0.001). This indicates that mode effects are present nationwide, including in the Midwest, Northeast, South, and West. The consistency of these findings across regions suggests that geographic variation does not fully explain differences in response rates, and that survey mode has a robust impact across diverse regional contexts.

```{r models-state4}
# State_4 logistic regression
full <- full %>%
  mutate(
    states_4 = factor(states_4)
  )

model_state4_main <- glm(
  Response ~ Mode + states_4,
  data = full %>% filter(!is.na(states_4), !is.na(Mode)),
  family = binomial()
)

model_state4_int <- glm(
  Response ~ Mode * states_4,
  data = full %>% filter(!is.na(states_4), !is.na(Mode)),
  family = binomial()
)

lrt_state4 <- anova(model_state4_main, model_state4_int, test = "Chisq")
lrt_state4
```

The interaction between mode and Census region (states_4) is not statistically significant (p = 0.76), indicating that the effect of survey mode on response rates is consistent across regions.

```{r summary-state4}
summary(model_state4_main)
summary(model_state4_int)
```

---

# Part 3: Employment Status (Ruisi)

```{r recode-employment}
# Recode Employment_status
# Missing values are kept as NA and excluded only
# when analyzing this subgroup variable
full <- full %>%
  mutate(
    Employment_recode = case_when(
      emp_code == "01" ~ "Employed full time",
      emp_code == "02" ~ "Employed part time",
      TRUE ~ NA_character_
    )
  )
```

```{r rr-employment}
# Response rate by Employment_status subgroup
# Exclude missing Employment_status only here
rr_emp <- full %>%
  filter(!is.na(Employment_recode), !is.na(Mode)) %>%
  group_by(Employment_recode, Mode) %>%
  summarise(
    n = n(),
    r = sum(Response, na.rm = TRUE),
    RR = r / n,
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = Mode,
    values_from = c(n, r, RR)
  ) %>%
  mutate(
    Delta = RR_Mail - RR_Web
  )
```

```{r chisq-employment}
# Chi-square tests within each subgroup
chisq_emp <- full %>%
  filter(!is.na(Employment_recode), !is.na(Mode)) %>%
  group_by(Employment_recode) %>%
  group_modify(~{
    tab <- table(.x$Mode, .x$Response)
    test <- chisq.test(tab, correct = FALSE)
    tibble(
      chi_square = unname(test$statistic),
      df = unname(test$parameter),
      p_value = test$p.value
    )
  })
```

```{r emp-table}
emp_table <- full %>%
  filter(!is.na(Employment_recode), !is.na(Mode)) %>%
  count(Employment_recode, Mode, Response) %>%
  pivot_wider(
    names_from = c(Mode, Response),
    values_from = n,
    values_fill = 0
  ) %>%
  rename(
    Mail_Nonresponse = Mail_0,
    Mail_Response = Mail_1,
    Web_Nonresponse = Web_0,
    Web_Response = Web_1
  )

rr_emp
chisq_emp
emp_table
```

```{r logistic-setup-employment}
# Set factor reference levels
# Web is the reference group, so the Mode coefficient is interpreted as Mail vs. Web
full <- full %>%
  mutate(
    Employment_recode = factor(
      Employment_recode,
      levels = c("Employed full time", "Employed part time")
    )
  )
```

```{r models-employment}
# Employment model without interaction
model_emp_main <- glm(
  Response ~ Mode + Employment_recode,
  data = full %>% filter(!is.na(Employment_recode), !is.na(Mode)),
  family = binomial()
)

# Employment model with interaction
model_emp_int <- glm(
  Response ~ Mode * Employment_recode,
  data = full %>% filter(!is.na(Employment_recode), !is.na(Mode)),
  family = binomial()
)
```

```{r lrt-employment}
# Likelihood ratio tests for interaction
lrt_emp <- anova(model_emp_main, model_emp_int, test = "Chisq")
lrt_emp
```

```{r summary-employment}
summary(model_emp_int)
```

```{r or-employment}
# Odds ratios and 95% CIs
emp_or <- tidy(model_emp_int, conf.int = TRUE, exponentiate = TRUE)
emp_or
```

```{r predicted-prob-employment}
# Predicted probabilities by mode and subgroup
emp_pred <- expand.grid(
  Mode = levels(full$Mode),
  Employment_recode = levels(full$Employment_recode)
) %>%
  mutate(
    Predicted_Probability = predict(
      model_emp_int,
      newdata = .,
      type = "response"
    )
  ) %>%
  pivot_wider(
    names_from = Mode,
    values_from = Predicted_Probability
  ) %>%
  mutate(
    Difference_Mail_minus_Web = Mail - Web
  )

emp_pred
```

```{r predicted-logit-employment}
# Predicted log-odds
emp_logit <- expand.grid(
  Mode = levels(full$Mode),
  Employment_recode = levels(full$Employment_recode)
) %>%
  mutate(
    Predicted_Logit = predict(
      model_emp_int,
      newdata = .,
      type = "link"
    )
  )

emp_logit
```

---

# Part 4: Job Function (Ruisi)

```{r recode-function}
# Recode Function
full <- full %>%
  mutate(
    Function_recode = case_when(
      func_code %in% c("0100", "0400", "0800", "0802", "0803") ~ "Practice and Supervision",
      func_code %in% c("0200", "0201", "0203", "0204", "0205", "0300") ~ "Academic and Research",
      func_code %in% c("0600", "0601", "0602", "0603", "0604", "0700", "0701", "0703", "0704") ~ "Administration and Leadership",
      func_code %in% c("0900", "0901") ~ "Other",
      TRUE ~ NA_character_
    )
  )
```

```{r rr-function}
# Response rate by Function subgroup
# Exclude missing Function only here
rr_func <- full %>%
  filter(!is.na(Function_recode), !is.na(Mode)) %>%
  group_by(Function_recode, Mode) %>%
  summarise(
    n = n(),
    r = sum(Response, na.rm = TRUE),
    RR = r / n,
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = Mode,
    values_from = c(n, r, RR)
  ) %>%
  mutate(
    Delta = RR_Mail - RR_Web
  )
```

```{r chisq-function}
chisq_func <- full %>%
  filter(!is.na(Function_recode), !is.na(Mode)) %>%
  group_by(Function_recode) %>%
  group_modify(~{
    tab <- table(.x$Mode, .x$Response)
    test <- chisq.test(tab, correct = FALSE)
    tibble(
      chi_square = unname(test$statistic),
      df = unname(test$parameter),
      p_value = test$p.value
    )
  })
```

```{r func-table}
func_table <- full %>%
  filter(!is.na(Function_recode), !is.na(Mode)) %>%
  count(Function_recode, Mode, Response) %>%
  pivot_wider(
    names_from = c(Mode, Response),
    values_from = n,
    values_fill = 0
  ) %>%
  rename(
    Mail_Nonresponse = Mail_0,
    Mail_Response = Mail_1,
    Web_Nonresponse = Web_0,
    Web_Response = Web_1
  )

rr_func
chisq_func
func_table
```

```{r logistic-setup-function}
full <- full %>%
  mutate(
    Function_recode = factor(
      Function_recode,
      levels = c(
        "Practice and Supervision",
        "Administration and Leadership",
        "Other",
        "Academic and Research"
      )
    )
  )
```

```{r models-function}
# Function model without interaction
model_func_main <- glm(
  Response ~ Mode + Function_recode,
  data = full %>% filter(!is.na(Function_recode), !is.na(Mode)),
  family = binomial()
)

# Function model with interaction
model_func_int <- glm(
  Response ~ Mode * Function_recode,
  data = full %>% filter(!is.na(Function_recode), !is.na(Mode)),
  family = binomial()
)
```

```{r lrt-function}
lrt_func <- anova(model_func_main, model_func_int, test = "Chisq")
lrt_func
```

```{r summary-function}
summary(model_func_int)
```

```{r or-function}
# Odds ratios and 95% CIs
func_or <- tidy(model_func_int, conf.int = TRUE, exponentiate = TRUE)
func_or
```

```{r predicted-prob-function}
# Predicted probabilities by mode and subgroup
func_pred <- expand.grid(
  Mode = levels(full$Mode),
  Function_recode = levels(full$Function_recode)
) %>%
  mutate(
    Predicted_Probability = predict(
      model_func_int,
      newdata = .,
      type = "response"
    )
  ) %>%
  pivot_wider(
    names_from = Mode,
    values_from = Predicted_Probability
  ) %>%
  mutate(
    Difference_Mail_minus_Web = Mail - Web
  )

func_pred
```

```{r predicted-logit-function}
# Predicted log-odds
func_logit <- expand.grid(
  Mode = levels(full$Mode),
  Function_recode = levels(full$Function_recode)
) %>%
  mutate(
    Predicted_Logit = predict(
      model_func_int,
      newdata = .,
      type = "link"
    )
  )

func_logit
```
