---
title: "Methods and Analysis"
format:
  pdf:
    documentclass: scrartcl
    fontsize: 11pt
    geometry: margin=1in
    linestretch: 1.1
    number-sections: true
    toc: false
    fig-pos: "H"
    keep-tex: false
execute:
  echo: false
  warning: false
  message: false
---

## 2 2024 SLP Schools Survey (Updated Analysis)

### 2.1 Brief Introduction

The 2024 SLP Schools Survey employs a mixed-mode design comparing postal mail and SurveyMonkey to evaluate differences in response behavior across key respondent characteristics. Mode is a central design feature because it directly affects accessibility, burden, and engagement. This section evaluates overall mode effects, subgroup variation, and whether differential response leads to meaningful composition bias.

### 2.2 Data Source

The 2024 SLP Schools Survey was administered to a sample of 15,000 SLPs working in school settings. After removing ineligible cases, 14,673 units remained eligible for analysis.

A total of 3,751 completed responses were identified after correcting the response definition, yielding an overall eligibility-adjusted response rate of:

$$
RR = 25.6\%
$$

The survey includes the following variables:

-   Mode (Postal Mail vs SurveyMonkey)
-   Employment status
-   Facility type
-   Age
-   State (recoded into Census regions)

### 2.3 Methods

#### Response Rate Comparison

Response rates were computed as:

$$
RR = \frac{\text{Completed responses}}{\text{Eligible sample}}
$$

Rates were compared across mode, employment, age, facility, and region.

#### Chi-Square Tests

Pearson chi-square tests were used to assess whether response differs across groups:

$$
\chi^2 = \sum \frac{(O - E)^2}{E}
$$

#### Logistic Regression

To evaluate adjusted effects, the following model was estimated:

$$
\log\left(\frac{p}{1-p}\right)
=
\beta_0 + \beta_1 \text{Mode} + \beta_2 \text{Employment} + \beta_3 \text{Facility} + \beta_4 \text{Age} + \beta_5 \text{Region}
$$

### 2.4 Results

#### Mode Effect

| Mode           | Response Rate |
|----------------|---------------|
| Postal Mail    | 30.0%         |
| SurveyMonkey   | 21.2%         |
| Δ (Mail − Web) | +8.8 pp       |

This indicates a strong and statistically significant mail advantage.

#### Age

-   \<35: 23.8%
-   35–44: 25.4%
-   45–54: 27.3%
-   55–64: 30.9%
-   65+: 24.4%

Response increases with age, peaking at 55–64.

#### Employment

-   Full-time: 25.8%
-   Part-time: 23.5%

Differences are statistically significant but small.

#### Facility Type

Response varies descriptively but is not statistically significant after adjustment. Facility type does not independently predict response.

#### Geographic (Region)

| Region    | Response Rate |
|-----------|---------------|
| Midwest   | 27.8%         |
| West      | 27.2%         |
| South     | 24.3%         |
| Northeast | 23.5%         |

Regional differences are modest. The Mode × Region interaction is not significant, indicating a consistent mail advantage across regions.

#### Composition Bias

-   Midwest: +2.1 pp
-   Northeast: −1.8 pp
-   South: −1.6 pp
-   West: +1.4 pp

Differences are small, indicating limited observable nonresponse bias.

### 2.5 Interpretation

Mode is the dominant driver of response, with postal mail significantly outperforming SurveyMonkey. Age is the most important respondent-level predictor, while employment and facility type play smaller roles.

Geographic differences exist but are modest and do not materially affect representativeness. Overall, response variation does not lead to severe composition distortion.

### 2.6 Key Visuals

#### Figure 1: Response Rate by Mode

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width=10, fig.height=4}
library(tidyverse)
library(readxl)
library(scales)

# ============================================================
# PART A: SLP 2024 SURVEY VISUALS
# ============================================================

# ------------------------------------------------------------
# 1. Read and clean SLP 2024 files
# ------------------------------------------------------------
resp <- read_excel("/Users/sagnikchakravarty/Desktop/UMD_College_Work/Semester 4/Survey Design/2024 SLPs Schools Respondent File for Capstone.xlsx")
full <- read_excel("/Users/sagnikchakravarty/Desktop/UMD_College_Work/Semester 4/Survey Design/2024 SLPs Schools Full Sample for Capstone.xlsx")

resp2 <- resp %>%
  transmute(
    new_id = as.integer(NewID),
    mode = as.integer(mode)
  )

full2 <- full %>%
  transmute(
    new_id = as.integer(NewID),
    mode = as.integer(Mode),
    state_code = toupper(as.character(State)),
    response_code = as.character(`Repsonse Code`)
  ) %>%
  mutate(
    mode_label = case_when(
      mode == 1 ~ "Postal Mail",
      mode == 2 ~ "SurveyMonkey",
      TRUE ~ NA_character_
    )
  )

state_to_region <- tibble(
  state_code = c(
    "CT","ME","MA","NH","RI","VT","NJ","NY","PA",
    "IL","IN","MI","OH","WI","IA","KS","MN","MO","NE","ND","SD",
    "DE","FL","GA","MD","NC","SC","VA","DC","WV","AL","KY","MS","TN","AR","LA","OK","TX",
    "AZ","CO","ID","MT","NV","NM","UT","WY","AK","CA","HI","OR","WA"
  ),
  region = c(
    rep("Northeast", 9),
    rep("Midwest", 12),
    rep("South", 17),
    rep("West", 13)
  )
)

postal_ineligible <- c(
  "2 - Bad Address",
  "4 - Ineligible: Deceased",
  "5 - Ineligible: Retired",
  "6 - Ineligible: Wrong Facility, etc.",
  "7 - Ineligible: Not working in the professions",
  "8 - Ineligible: Other Reason"
)

slp_dat <- full2 %>%
  left_join(state_to_region, by = "state_code") %>%
  mutate(
    responded_from_frame = case_when(
      mode == 1 & response_code == "1" ~ 1L,
      TRUE ~ 0L
    ),
    responded_from_respfile = if_else(new_id %in% resp2$new_id, 1L, 0L),
    responded = case_when(
      mode == 1 ~ if_else(responded_from_frame == 1L | responded_from_respfile == 1L, 1L, 0L),
      mode == 2 ~ responded_from_respfile,
      TRUE ~ 0L
    ),
    eligible = case_when(
      mode == 1 & response_code %in% postal_ineligible ~ 0L,
      mode == 2 & response_code %in% c("Bounced", "Opted out") ~ 0L,
      TRUE ~ 1L
    )
  )

# ------------------------------------------------------------
# 2. Figure 1: Response Rate by Mode
# ------------------------------------------------------------
rr_mode <- slp_dat %>%
  filter(!is.na(mode_label)) %>%
  group_by(mode_label) %>%
  summarize(
    eligible_n = sum(eligible, na.rm = TRUE),
    complete_n = sum(responded, na.rm = TRUE),
    response_rate = complete_n / eligible_n,
    .groups = "drop"
  )

fig1 <- ggplot(rr_mode, aes(x = mode_label, y = response_rate)) +
  geom_col(width = 0.65, fill = "steelblue") +
  geom_text(
    aes(label = percent(response_rate, accuracy = 0.1)),
    vjust = -0.5,
    size = 5
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, max(rr_mode$response_rate) + 0.06)
  ) +
  labs(
    title = "Response Rate by Survey Mode",
    x = NULL,
    y = "Response Rate"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(face = "bold")
  )

print(fig1)
```

#### Figure 2: Response Rate by Region and Mode

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width=10, fig.height=4}
# ------------------------------------------------------------
# 3. Figure 2: Response Rate by Region and Mode
# ------------------------------------------------------------
rr_region_mode <- slp_dat %>%
  filter(!is.na(region), !is.na(mode_label), eligible == 1) %>%
  group_by(region, mode_label) %>%
  summarize(
    eligible_n = n(),
    complete_n = sum(responded, na.rm = TRUE),
    response_rate = complete_n / eligible_n,
    .groups = "drop"
  ) %>%
  mutate(region = factor(region, levels = c("Northeast", "Midwest", "South", "West")))

fig2 <- ggplot(rr_region_mode, aes(x = region, y = response_rate, fill = mode_label)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  geom_text(
    aes(label = percent(response_rate, accuracy = 0.1)),
    position = position_dodge(width = 0.75),
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0, max(rr_region_mode$response_rate) + 0.07)
  ) +
  labs(
    title = "Response Rate by Region and Survey Mode",
    x = "Region",
    y = "Response Rate",
    fill = "Mode"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "top"
  )

print(fig2)

```


## 3. Postcard Follow-up Analysis (Q3)

### 3.1 Brief Introduction

This section evaluates the effect of follow-up postcards on response rates. Because postcards are sent only to nonrespondents, naïve comparisons introduce treatment-by-indication bias. A sequential framework is used to estimate unbiased effects.

### 3.2 Methods

Define:

-   $RR_1$: Response before first postcard
-   $RR_2$: Response among no-card group after first stage
-   $RR_3$: Response among one-card group after first stage
-   $RR_4$: Response among one-card group after second stage
-   $RR_5$: Response among two-card group after second stage

#### No postcard

$$
P(R|card=0) = RR_1 + (1 - RR_1)RR_2
$$

#### One postcard

$$
P(R|card=1) = RR_1 + (1 - RR_1)\left[RR_3 + (1 - RR_3)RR_4\right]
$$

#### Two postcards

$$
P(R|card=2) = RR_1 + (1 - RR_1)\left[RR_3 + (1 - RR_3)RR_5\right]
$$

Bootstrap resampling is used to estimate uncertainty.

### 3.3 Results

| Scenario      | Response Rate |
|---------------|---------------|
| No postcard   | 24.3%         |
| One postcard  | 26.5%         |
| Two postcards | 27.0%         |

#### Effects

-   First postcard: +2.2 percentage points
-   Second postcard: +0.5 percentage points

### 3.4 Interpretation

The first postcard produces a meaningful increase in response, while the second postcard adds very little.

This reflects diminishing returns: early nonrespondents are responsive to follow-up, while late nonrespondents are more resistant.

The corrected estimates are smaller than naïve comparisons because they properly account for selection into treatment groups.

### 3.5 Recommendation

A single postcard is effective and cost-efficient. A second postcard provides minimal additional benefit unless the design is improved.

### 3.6 Key Visual

#### Figure 3: Incremental Effect of Postcards

```{r warning=FALSE, message=FALSE, echo=FALSE, fig.width=10, fig.height=4}
# ============================================================
# PART B: Q3 POSTCARD VISUAL
# ============================================================

library(lubridate)
library(stringr)

# ------------------------------------------------------------
# 4. Read and clean postcard experiment files
# ------------------------------------------------------------
file_path <- "/Users/sagnikchakravarty/Desktop/Postcard Experiment IDs for Capstone.xlsx"

full_sample <- read_excel(file_path, sheet = "Full Sample")
pc1_ids <- read_excel(file_path, sheet = "Postcard 1 Nonresponders")
pc2_ids <- read_excel(file_path, sheet = "Postcard 2 Nonresponders")
respondents <- read_excel(file_path, sheet = "Survey Respondents with Date")

clean_names <- function(x) {
  x <- str_trim(x)
  x <- str_replace_all(x, "\\s+", "_")
  x
}

names(full_sample)[1] <- "ID"
names(pc1_ids)[1] <- "ID"
names(pc2_ids)[1] <- "ID"
names(respondents)[1] <- "ID"

names(full_sample) <- clean_names(names(full_sample))
names(pc1_ids) <- clean_names(names(pc1_ids))
names(pc2_ids) <- clean_names(names(pc2_ids))
names(respondents) <- clean_names(names(respondents))

end_candidates <- c("EndDate", "End_Date", "End", "Submit_Date", "Submitted_Date", "Submission_Date")
end_col <- end_candidates[end_candidates %in% names(respondents)][1]

id_candidates <- c("ID", "Id", "Respondent_ID", "RespondentId", "Anonymized_ID", "AnonymizedID")
id_col <- id_candidates[id_candidates %in% names(respondents)][1]

respondents <- respondents %>%
  rename(
    ID = all_of(id_col),
    EndDate_raw = all_of(end_col)
  )

parse_date_robust <- function(x) {
  if (inherits(x, "Date")) return(x)
  if (inherits(x, "POSIXct") || inherits(x, "POSIXt")) return(as.Date(x))
  if (is.numeric(x)) return(as.Date(x, origin = "1899-12-30"))
  x1 <- suppressWarnings(as.Date(x))
  x2 <- suppressWarnings(mdy(x))
  x3 <- suppressWarnings(ymd(x))
  x4 <- suppressWarnings(dmy(x))
  coalesce(x1, x2, x3, x4)
}

respondents <- respondents %>%
  mutate(EndDate = parse_date_robust(EndDate_raw)) %>%
  select(ID, EndDate)

# ------------------------------------------------------------
# 5. Compute postcard scenario estimates
# ------------------------------------------------------------
t1 <- as.Date("2025-03-13")
t2 <- as.Date("2025-04-10")

full_ids <- full_sample %>% distinct(ID)
pc1_ids <- pc1_ids %>% distinct(ID)
pc2_ids <- pc2_ids %>% distinct(ID)

no_card_ids <- full_ids %>% filter(!ID %in% pc1_ids$ID)
first_card_ids <- pc1_ids
one_card_ids <- pc1_ids %>% filter(!ID %in% pc2_ids$ID)
two_card_ids <- pc2_ids

all_dat <- full_ids %>%
  left_join(respondents, by = "ID") %>%
  mutate(
    responded_by_t1 = !is.na(EndDate) & EndDate <= t1,
    responded_t1_t2 = !is.na(EndDate) & EndDate > t1 & EndDate <= t2,
    responded_after_t2 = !is.na(EndDate) & EndDate > t2
  )

no_card_dat <- no_card_ids %>% left_join(all_dat, by = "ID")
first_card_dat <- first_card_ids %>% left_join(all_dat, by = "ID")
one_card_dat <- one_card_ids %>% left_join(all_dat, by = "ID")
two_card_dat <- two_card_ids %>% left_join(all_dat, by = "ID")

RR1 <- mean(all_dat$responded_by_t1)

no_card_t1_nonresp <- no_card_dat %>% filter(!responded_by_t1)
RR2 <- mean((!is.na(no_card_t1_nonresp$EndDate)) & no_card_t1_nonresp$EndDate > t1)

first_card_t1_nonresp <- first_card_dat %>% filter(!responded_by_t1)
RR3 <- mean(first_card_t1_nonresp$responded_t1_t2)

one_card_t2_nonresp <- one_card_dat %>% filter(!responded_by_t1, !responded_t1_t2)
RR4 <- mean(one_card_t2_nonresp$responded_after_t2)

two_card_t2_nonresp <- two_card_dat %>% filter(!responded_by_t1, !responded_t1_t2)
RR5 <- mean(two_card_t2_nonresp$responded_after_t2)

P_card0 <- RR1 + (1 - RR1) * RR2
P_card1 <- RR1 + (1 - RR1) * (RR3 + (1 - RR3) * RR4)
P_card2 <- RR1 + (1 - RR1) * (RR3 + (1 - RR3) * RR5)

postcard_results <- tibble(
  postcards = factor(
    c("0 postcards", "1 postcard", "2 postcards"),
    levels = c("0 postcards", "1 postcard", "2 postcards")
  ),
  response_rate = c(P_card0, P_card1, P_card2)
)

# ------------------------------------------------------------
# 6. Figure 3: Incremental Effect of Postcards
# ------------------------------------------------------------
fig3 <- ggplot(postcard_results, aes(x = postcards, y = response_rate, group = 1)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_text(
    aes(label = percent(response_rate, accuracy = 0.1)),
    vjust = -0.8,
    size = 4.5
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    limits = c(0.12, max(postcard_results$response_rate) + 0.02)
  ) +
  labs(
    title = "Estimated Response Rate by Number of Postcards",
    x = "Intervention Scenario",
    y = "Estimated Response Rate"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold")
  )

print(fig3)
```
